Copulas and their potential for ecology

### 2019-10-17 Paper.Rmd
edits to Q2 sections and Master figure

For multivariate datasets, we performed the bivariate analysis
for all possible pairwise combinations of distinct locations. We carried out
pairwise bivariate analyses instead of trying to fit a multivariate copula,
for simplicity and because that approach was
sufficient to answer our research questions; but see the Discussion
for a few words on multivariate copulas. 
We present
the number of pairs for which a non-normal copula
was the AIC-best copula, and we characterize AIC differences between
the normal and AIC-best copulas across location pairs. 
We also computed the model-averaged lower- and
upper-tail statistics, and differences
between these, for each pair of locations, and we characterize
the distribution of these values across location pairs.

In addition to our model selection approach, we also used a
nonparametric approach, to provide greater 
confidence in our answers to Q1. We used three 
statistics which quantify the extent to which 
the normalized ranks $u_i$ and $v_i$ are 
related in any part of their distributions; we here describe the 
statistics, with some additional details in Appendix \ref{SM-nonparam_stats}.
The statistics are defined with positively associated
variables in mind. All our variables were positively associated
when they were significantly associated.
Given two bounds 
$0 \leq l_b < u_b \leq 1$, define the lines $u+v=2l_b$ and 
$u+v=2u_b$, which intersect the unit square (Fig. \ref{NonparamStatsFig}). 
Our statistics quantify the association
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and $u_b\leq 0.5$ gives information about 
association in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ gives information about association in 
the right parts. The first statistic, a partial Spearman correlation
\begin{equation}
\cor_{l_b,u_b}(u,v)=\frac{\sum 
   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}},\label{eq:partialspearman}
\end{equation}
\noindent is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
between the bounds. 
Here sample means and variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$.
Larger values of the partial Spearman correlation indicate stronger positive association.
The sum of 
$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
interval $(0,1)$) equals the standard Spearman correlation
as long as no points fall directly on the bounds.
We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
which has a similar qualitative
interpretation to $\cor_{l_{b},u_{b}}$.
Our third statistic, $\Dsq_{l_b,u_b}$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$.
Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, for which large values inidicate strong association
between the bounds, small values of 
$\Dsq_{l_b,u_b}$ indicate strong association. These statistics are not estimators of the 
tail dependence quantities
defined previously, but rather are conceptually similar measures of associations in
the tail portions of the distributions when appropriate values of $l_b$ and $u_b$ are
selected.

@@

For large datasets (large $n$),
we used $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we considered 
multiple bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$,
abbreviating $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"). Likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, $\Dsq_u=\Dsq_{0.5,1}$.
To test for asymmetry of dependence in upper and lower portions of distributions,
we used differences $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
for smaller datasets (note the opposite order
in the last of these); and $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ with $u_b$ close 
to $0$ for large datasets.
We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also 
Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}).

@@

For multivariate datasets, we calculated Spearman and Kendall correlations
and the statistics 
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
for all pairs of sampling locations, and we then computed means across the pairs
for each statistic. We used a spatial resampling scheme 
(Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals of these means.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models.
Code and data for this
project are archived at \url{www.github.com/sghosh89/BIVAN}.

@@

<!--bivariate example, model selection-->
For the soil C and N data (Data section, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.min(BivMS_res_Loecke$InfCritRes$AIC)]` was the 
minimum-AIC copula, with AIC `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)`, whereas the 
normal copula had a much higher AIC, `r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)`, answering the 
first part of Q1 for these data. Recall that AIC differences of $2$
or $3$ are considered meaningful and differences of $8$ or greater
are considered diffinitive, so that data provide essentially no support
for the higher-AIC model in that case [@burnham2003_modelselection].
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what a normal copula would give, namely, 0, helping to answer 
the second part 
Q1 for these data. The values also differed substantially 
from each other, helping answer the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).

@@

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets, with some details in the Appendices. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, row 2). 
The normal copula was a poor fit compared to the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
difference between these fits was not meaningfully large.
Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
Often, either lower- or upper-tail dependence statistics differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
helping to answer the second and third parts of Q1. Though again the model
selection results do not provide 
information on statistical significance of these differences, and are subject to the
caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
(Table \ref{tab_bivar_summary}, rows 5 and 6), nonparametric results 
(Table \ref{tab_bivar_summary}, row 10) showed that tail associations
deviated significantly from what would be expected from a normal-copula
null model, and were also significantly asymmetric, except for the Cedar Creek data.
See Figs \ref{SM-fig_bmr_birds_nonparam}-\ref{SM-fig_cc2000_nonparam} for analogues to 
Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
data, see Tables \ref{SM-tab_bmr_birds_fit}-\ref{SM-tab_cc2000fit} for analogues to Table \ref{tab_soilCNfit},
and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though non-significantly. Our 
generally affirmative (except for Cedar Creek), 
empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of 
the soil C and N and bird and mammal body mass versus metabolic rate
datasets, in the middle row of boxes.
We did the same analyses 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, biomass and Shannon's index were independent; for 1997-1999, results 
were similar to 2000.

@@

Asymmetry results were generally verified by nonparametric approaches,
with the exception of the methane-flux data. 
For instance, the $95\%$ confidence intervals of the
mean over pairs of sampling locations of the statistic $\cor_l-\cor_u$ were 
(`r unname(round(stat_aphid_ff$numericdf[9,c(3,4)],3))`)
for the leaf-curling plum aphid first flight data
(Table \ref{tab_multivar_summary}, row 9), 
indicating greater upper-tail dependence, and consistent with
the results of Table \ref{tab_multivar_summary}, row 8. 
For the *Ceratium furca* abundance data, confidence intervals were
(`r unname(round(stat_plankton_north_sea$numericdf[9,c(3,4)],3))`),
indicating greater lower-tail dependence, and again 
consistent with the results of Table \ref{tab_multivar_summary} (row 9).
For the methane-flux data, confidence intervals were 
(`r unname(round(stat_methane$numericdf[9,c(3,4)],3))`);
the asymmetry in tail dependence revealed by the model selection
results for the methane-flux data was apparently not strong enough to
also be revealed by the nonparametric analyses. 
Analogues to Table \ref{tab_count_resamp} are in Tables
\ref{SM-tab_ff_resamp}-\ref{SM-tab_methane_resamp}.
Our generally
affirmative, empirically based answer to Q1 for the multivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names
of those datasets, in the middle row of boxes.
We also carried out the same analyses for abundance and first-flight data for the 18 
aphid species for which we had data other than the green spruce and leaf-curling plum aphids,
as well as for the 21 plankton taxa for which we had abundance data other than
\textit{Ceratium furca} (results not shown); results supported the conclusion that 
non-normal copula structure and tail dependence, and asymmetry of tail associations, are 
common.

@@

Having demonstrated that non-normal copula structure and asymmetric tail associations are 
common in ecological variables, we 
addressed Q2 by exploring, using models, three possible mechanisms
by which these phenomena may arise. 
Models presented are initial explorations, only, of whether the proposed mechanisms have the 
potential to explain observed patterns. As such, simple models were used, 
and comprehensive explorations of
model parameter space and alternative model formulations were left for future work.

The first mechanism relates to the ideas in the Introduction about 
Liebig's law of the minimum, and to nonlinear influences of environmental
variables on ecological variables.
If an environmental variables influences an ecological variable 
disproportionately strongly in one of its tails, we explored whether the ecological
variable could then exhibit asymmetric tail dependence across space.
Let $E_i(t)$ be an ecological variable 
measured at location $i$ ($i=1,2$) at time $t$.
Assume the dynamics 
$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where 
the $\delta_i(t)$ are standard-normally distributed 
and independent across time and locations, $a=0.2$,
$b=0.1, -0.1, 0.5, -0.5$ in different simulations, and 
the $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
$\cov(\epsilon_1,\epsilon_2)=0.8$. Thus basic ecological dynamics follow the
very simple AR(1) formulation, influenced by a "regional" environmental factor which
is correlated across locations ($\epsilon$) and by an entirely local factor ($\delta$). 
The function $g$ describes how the environmental factor $\epsilon_i(t)$ influences
$E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise;
$g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
$g_1$ represents environmental effects that negatively impact populations,
and that occur only below a threshold of $\epsilon=0$; 
and $g_2$ represents effects that positively impact populations and 
that occur only above a threshold of $\epsilon=0$.
The values of $b$ selected provide a modicum of exploration of the potential
for ecological dynamics to influence how complex dependencies between
the $E_i(t)$ may arise: 
negative values of $b$ correspond to overcompensating dynamics 
and positive values to undercompensating dynamics; larger $|b|$
means slower return to equilibrium after a disturbance.

@@

Our second mechanism is an extension of the well known Moran
effect, and was summarized conceptually in the Introduction. 
We consider a linear model, as well as 
two parameterizations of a nonlinear population model which includes density dependence
of population growth. For the linear model, let $E_i(t)$ 
again be an ecological variable, $i=1,2$.
We use AR(1) dynamics, 
$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
The environmental noises $\epsilon_i(t)$ were 
standard-normal random variables that were independent for distinct times, $t$, but 
exhibited different kinds of 
dependence across locations in different simulations (see below).
The variable $E$ is very general, and could represent deviations
of a population density from a carrying capacity, deviations of
total plant community biomass from an average value,
flux of a biogeochemical variable such as methane, or other quantities.
The nonlinear population model was a stochastic, multi-habitat-patch version of the Ricer model, 
$P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$ for $i=1,2$, using $r=0.5$, 
$K=100$, $\epsilon_i(t)$
as above, and $\sigma=0.1$ or 
$\sigma=1$ in different simulations. 
When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
$K$, where the model equation can be well approximated by a linear equation,
and the nonlinearities of the model therefore have limited influence on dynamics. When
$\sigma=1$, model dynamics are strongly nonlinear because the stochastic 
component of the model causes populations to stray far from the carrying capacity, $K$. 
We refer to these as
the weak-noise and strong-noise cases, though the importance of the noise here
is that, when it is strong, it brings the nonlinearities of the model into play.

@@

The next mechanism we investigated is evolutionary, and pertains to bivariate 
trait data across species, e.g., our bird and mammal 
data. This mechanism is a hypothetical explanation for the bias toward right tail association
observed in those data (Fig. \ref{fig_biv_multi_raw_cop}F,G). 
The hypothesis is that asymmetric tail association occurs in evolutionary changes
in bivariate characters, and gives rise to asymmtric tail association between the 
two character values across extant species.
We simulated bivariate character evolution on an estimate of 
the phylogeny, taken from @GenoudIM2018, of $817$ mammal species.
The root character state and the change 
across each branch were randomly chosen from matrices of one million 
independent draws from bivariate distributions showing one of five distinct types of copula 
structure: 1) extreme or 2) moderate left-tail dependence,
3) symmetric tail dependence, or 4) moderate or 5) extreme
right-tail dependence (Appendix \ref{SM-evol_mech}). 
All distributions had standard-normal marginals and Spearman 
correlation $0.875$ between components, so our simulations assess 
the impact of copula structure only. For 
each of the five copulas, mammalian character
evolution was simulated 100 times. For each simulation, symmetry of tail dependence 
of the two characters across phylogeny tips was assessed using our nonparametric
statistics.
We hypothesized that cases 1 and 2 above would yield stronger left- than right-tail 
associations in tip characters, and cases 4 and 5 would yield the reverse. 
The simulator was written in Python and used version 4.4 of the DendroPy 
package [@DendroPy].

>>>>

For multivariate datasets, we performed the bivariate analysis
for all possible pairwise combinations of distinct locations. We carried out
pairwise bivariate analyses instead of trying to fit a multivariate copula,
for simplicity and because that approach was
sufficient to answer our research questions; but see the Discussion
for a few words on multivariate copulas. 
We present
the number of pairs for which a non-normal copula
was the AIC-best copula, and we characterize AIC differences between
the normal and AIC-best copulas across location pairs. 
We also computed the model-averaged lower- and
upper-tail statistics, and differences
between these, for each pair of locations, and we characterize
the distributions of these values across location pairs.

In addition to our model selection approach, we also used a
nonparametric approach, to provide greater 
confidence in our answers to Q1. We used three 
statistics which quantify the extent to which 
the normalized ranks $u_i$ and $v_i$ are 
related in any part of their distributions. We here describe the 
statistics, with some additional details in Appendix \ref{SM-nonparam_stats}.
The statistics are defined with positively associated
variables in mind. All our variables were positively associated
when they were significantly associated.
Given two bounds 
$0 \leq l_b < u_b \leq 1$, define the lines $u+v=2l_b$ and 
$u+v=2u_b$, which intersect the unit square (Fig. \ref{NonparamStatsFig}). 
Our statistics quantify the association
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and $u_b\leq 0.5$ gives information about 
association in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ gives information about association in 
the right parts. The first statistic, a partial Spearman correlation,
\begin{equation}
\cor_{l_b,u_b}(u,v)=\frac{\sum 
   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}},\label{eq:partialspearman}
\end{equation}
\noindent is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
between the bounds. 
Here sample means and variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$.
Larger values of the partial Spearman correlation indicate stronger positive association.
The sum of 
$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
interval $(0,1)$) equals the standard Spearman correlation
as long as no points fall exactly on the bounds.
We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
which has a similar qualitative
interpretation to $\cor_{l_{b},u_{b}}$.
Our third statistic, $\Dsq_{l_b,u_b}$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$.
Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, for which large values indicate strong association
between the bounds, small values of 
$\Dsq_{l_b,u_b}$ indicate strong association. These statistics are not estimators of the 
tail dependence quantities
defined previously, but rather are conceptually similar measures of associations in
the tail portions of the distributions when appropriate values of $l_b$ and $u_b$ are
selected.

@@

For large datasets (large $n$),
we used $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we considered 
multiple bands $(l_b,u_b)$ to understand how association
varies across different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$,
abbreviating $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"). Likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, $\Dsq_u=\Dsq_{0.5,1}$.
To test for asymmetry of dependence in upper and lower portions of distributions,
we used differences $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
for smaller datasets (note the opposite order
in the last of these); and $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ with $u_b$ close 
to $0$ for large datasets.
We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also 
Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}).

@@

For multivariate datasets, we calculated Spearman and Kendall correlations
and the statistics 
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
for all pairs of sampling locations, and we then computed means across the pairs
for each statistic. We used a spatial resampling scheme 
(Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals of these means.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method. The same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models.
Code and data for this
project are archived at \url{www.github.com/sghosh89/BIVAN}.

@@

<!--bivariate example, model selection-->
For the soil C and N data (Data section, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). Non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.min(BivMS_res_Loecke$InfCritRes$AIC)]` was the 
minimum-AIC copula, with AIC `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)`, whereas the 
normal copula had a much higher AIC, `r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)`, answering the 
first part of Q1 for these data. Recall that AIC differences of $2$
or $3$ are considered meaningful and differences of $8$ or greater
are considered diffinitive, so that data provide essentially no support
for the higher-AIC model in that case [@burnham2003_modelselection].
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what a normal copula would give, namely, 0, helping to answer 
the second part 
Q1 for these data. The values also differed substantially 
from each other, helping answer the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).

@@

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets, with some details in the Appendices. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, row 2). 
The normal copula was a poor fit compared to the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
difference between these fits was marginal.
Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
Often, either lower- or upper-tail dependence statistics differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
helping to answer the second and third parts of Q1. Though again the model
selection results do not provide 
information on statistical significance of these differences, and are subject to the
caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
(Table \ref{tab_bivar_summary}, rows 5 and 6), nonparametric results 
(Table \ref{tab_bivar_summary}, row 10) showed that tail associations
deviated significantly from what would be expected from a normal-copula
null model, and were also significantly asymmetric, except for the Cedar Creek data.
See Figs \ref{SM-fig_bmr_birds_nonparam}-\ref{SM-fig_cc2000_nonparam} for analogues to 
Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
data, see Tables \ref{SM-tab_bmr_birds_fit}-\ref{SM-tab_cc2000fit} for analogues to Table \ref{tab_soilCNfit},
and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though non-significantly. Our 
generally affirmative (except for Cedar Creek), 
empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of 
the soil C and N and bird and mammal body mass versus metabolic rate
datasets, in the middle row of boxes.
We did the same analyses 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, independence of biomass and Shannon's index could not be rejected; 
for 1997-1999, results were similar to 2000.

@@

Asymmetry results were generally verified by nonparametric approaches,
with the exception of the methane-flux data. 
For instance, the $95\%$ confidence intervals of the
mean over pairs of sampling locations of the statistic $\cor_l-\cor_u$ were 
(`r unname(round(stat_aphid_ff$numericdf[9,c(3,4)],3))`)
for the leaf-curling plum aphid first flight data
(Table \ref{tab_multivar_summary}, row 9), 
indicating greater upper-tail dependence, and consistent with
the results of Table \ref{tab_multivar_summary}, row 8. 
For the *Ceratium furca* abundance data, confidence intervals were
(`r unname(round(stat_plankton_north_sea$numericdf[9,c(3,4)],3))`),
indicating greater lower-tail dependence, and again 
consistent with the results of Table \ref{tab_multivar_summary}, row 8.
For the methane-flux data, confidence intervals were 
(`r unname(round(stat_methane$numericdf[9,c(3,4)],3))`);
the asymmetry in tail dependence revealed by the model selection
results for the methane-flux data was apparently not strong enough to
also be revealed by the nonparametric analyses. 
Analogues to Table \ref{tab_count_resamp} are in Tables
\ref{SM-tab_ff_resamp}-\ref{SM-tab_methane_resamp}.
Our generally
affirmative, empirically based answer to Q1 for the multivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names
of those datasets, in the middle row of boxes.
We also carried out the same analyses for abundance and first-flight data for the 18 
aphid species for which we had data other than the green spruce and leaf-curling plum aphids,
as well as for the 21 plankton taxa for which we had abundance data other than
\textit{Ceratium furca} (results not shown). Results supported the conclusion that 
non-normal copula structure and tail dependence, and asymmetry of tail associations, are 
common.

@@

Having demonstrated that non-normal copula structure and asymmetric tail associations are 
addressed Q2 by exploring, using models, three possible mechanisms
by which these phenomena may arise. 
Models presented are initial explorations, only, of whether the proposed mechanisms have the 
potential to explain observed patterns. As such, simple models were used. 
Comprehensive explorations of
model parameter space and alternative model formulations were left for future work.

The first mechanism relates to the ideas in the Introduction about 
Liebig's law of the minimum, and to nonlinear influences of environmental
variables on ecological variables.
If an environmental variables influences an ecological variable 
disproportionately in one of its tails, we explored whether the ecological
variable could then exhibit asymmetric tail associations across space.
Let $E_i(t)$ be an ecological variable 
measured at location $i$ ($i=1,2$) at time $t$.
Assume the dynamics 
$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where 
the $\delta_i(t)$ are standard-normally distributed 
and independent across time and locations, $a=0.2$,
$b=0.1, -0.1, 0.5, -0.5$ in different simulations, and 
the $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
$\cov(\epsilon_1,\epsilon_2)=0.8$. Thus basic ecological dynamics follow the
very simple AR(1) formulation, influenced by a "regional" environmental factor which
is correlated across locations ($\epsilon$) and by an entirely local factor ($\delta$). 
The function $g$ describes how the environmental factor $\epsilon_i(t)$ influences
$E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise;
$g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
$g_1$ represents environmental effects that negatively impact populations,
and that occur only below a threshold of $\epsilon=0$; 
and $g_2$ represents effects that positively impact populations and 
that occur only above a threshold of $\epsilon=0$.
The values of $b$ selected provide a modicum of exploration of the potential
for ecological dynamics to also influence how complex dependencies between
the $E_i(t)$ may arise: 
negative values of $b$ correspond to overcompensating dynamics 
and positive values to undercompensating dynamics; larger $|b|$
means slower return to equilibrium after a disturbance.

@@

Our second mechanism is an extension of the well known Moran
effect, and was summarized conceptually in the Introduction. 
We consider a linear model, as well as 
two parameterizations of a nonlinear population model which includes density dependence
of population growth. For the linear model, let $E_i(t)$ 
again be an ecological variable, $i=1,2$.
We use AR(1) dynamics, 
$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
The environmental noises $\epsilon_i(t)$ were 
standard-normal random variables that were independent for distinct times, $t$, but 
exhibited different kinds of 
dependence across locations in different simulations (see below).
The variable $E$ is very general, and could represent deviations
of a population density from a carrying capacity, deviations of
total plant community biomass from an average value,
flux of a biogeochemical variable such as methane, or other quantities.
The nonlinear population model was a stochastic, multi-habitat-patch version 
of the Ricker model, 
$P_i(t+1)=P_i(t)\exp \left[r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right]$ for $i=1,2$, using $r=0.5$, 
$K=100$, $\epsilon_i(t)$
as above, and $\sigma=0.1$ or 
$\sigma=1$ in different simulations. 
When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
$K$, where the model equation can be well approximated by a linear equation,
and the nonlinearities of the model therefore have limited influence on dynamics. When
$\sigma=1$, model dynamics are strongly nonlinear because the stochastic 
component of the model causes populations to stray far from the carrying capacity, $K$. 
We refer to these as
the weak-noise and strong-noise cases, though the importance of the noise here
is that, when it is strong, it brings the nonlinearities of the model into play.

@@

The next mechanism we investigated is evolutionary, and pertains to bivariate 
trait data across species, e.g., our bird and mammal 
data. This mechanism is a hypothetical explanation for the bias toward right tail association
observed in those data (Fig. \ref{fig_biv_multi_raw_cop}F,G). 
The hypothesis is that asymmetric tail association occurs in evolutionary changes
in bivariate characters, and gives rise to asymmtric tail association between the 
two character values across extant species.
We simulated bivariate character evolution on an estimate of 
the phylogeny, taken from @GenoudIM2018, of $817$ mammal species.
The root character state and the change 
across each branch were randomly chosen from matrices of one million 
independent draws from bivariate distributions showing one of five distinct types of copula 
structure: 1) extreme or 2) moderate left-tail dependence,
3) symmetric tail dependence, or 4) moderate or 5) extreme
right-tail dependence (Appendix \ref{SM-evol_mech}). 
All distributions had standard-normal marginals and Spearman 
correlation $0.875$ between components, so our simulations assess 
the impact of copula structure only. For 
each of the five copulas, mammalian character
evolution was simulated 100 times. For each simulation, symmetry of tail associations 
of the two characters across phylogeny tips was assessed using our nonparametric
statistics.
We hypothesized that cases 1 and 2 above would yield stronger left- than right-tail 
associations in tip characters, and cases 4 and 5 would yield the reverse. 
The simulator was written in Python and used version 4.4 of the DendroPy 
package [@DendroPy].
